---
title: "Practical Machine Learning - Course Project"
output: html_document
---
Objective : Analyze the data coming from accelerometers present in the activity bands (belt, forearm, arm, and dumbell) of the 6 participants in the study and predict the manner in which they did the exercise. (They were asked to perform barbell lifts correctly and incorrectly in 5 different ways)

Dataset : The data for this project came from the source http://groupware.les.inf.puc-rio.br/har

Training dataset : 19622 rows * 160 cols (including response column ‘classe’)

Holdout sample: 20 records * 160 cols (response column has been replaced by a dummy/surrogate key)

Load the Training & Test/Holdout data files and do a basic inspection of the fields, data types, values etc.
```{r}
setwd("/home/vulcan/R/Coursera/PracticalMachineLearning-JohnsHopkins")
sensorTrain <- read.csv("pml-training.csv", header = TRUE, sep = ",")
sensorTest <- read.csv("pml-testing.csv", header = TRUE, sep = ",")

head(sensorTrain,10)
str(sensorTrain)
nrow(sensorTrain)
ncol(sensorTrain)
```
  
### Several columns seem to have *NA*,*NULL*, and other invalid values like *‘#DIV/0!’*

Remove all the columns which have NA and NULL in more than 40% of records
```{r}
cleanPredSensorTrain <- sensorTrain
which(as.numeric(colSums(is.na(cleanPredSensorTrain)))/nrow(cleanPredSensorTrain) > .4)
cleanPredSensorTrain <- cleanPredSensorTrain[,-which(as.numeric(colSums(is.na(cleanPredSensorTrain)))/nrow(cleanPredSensorTrain) > .4)]
cleanPredSensorTrain <- cleanPredSensorTrain[,-which(colSums(cleanPredSensorTrain == '')/nrow(cleanPredSensorTrain) > .4)]
str(cleanPredSensorTrain)
```
  
### Split this dataset for model estimation & tuning (80/20)
```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(33833)
trainIndex <- createDataPartition(cleanPredSensorTrain$classe, p = .80,list=FALSE,times=1)
cleanTrain <- cleanPredSensorTrain[trainIndex,]
cleanTest  <- cleanPredSensorTrain[-trainIndex,]
```
  
### Check for predictors with near zero variance. (excluding User Names, Response and time related fields)
```{r}
nearZeroVar(cleanTrain[,-c(1,2,3,4,5,60)])
```
  
### Removing User Names, Response,time related fields, and anything that is returned from above (new_window)
```{r}
cleanTrain <- cleanTrain[,-c(1,2,3,4,5,6)]
```
  
### Look for Between-Predictor correlations and plot the correlation matrix.
```{r}
correlations <- cor(cleanTrain[,-54])
```

```{r,echo=FALSE}
library(corrplot)
corrplot(correlations,order="hclust")
```
  
### Eliminate the predictors which have a between predictor correlation beyond a certain threshold.

```{r}
highCorr <- findCorrelation(correlations,cutoff = .75)
highCorr
length(highCorr)
cleanTrain <- cleanTrain[,-highCorr]
names(cleanTrain)
```

### Model fit 
  
1. Random Forest CrossValidation - Setting up for K fold CV. K = 5

```{r}
#set.seed(33833)
#train_control <- trainControl(method="cv", number=5)
#finalM1 <- train(classe ~ ., method = "rf",trControl=train_control, data = cleanTrain,prox = TRUE)
#save(finalM1,file="finalM1.Rda")
load("finalM1.Rda")
finalM1
confusionMatrix(finalM1)
 varImp(finalM1, scale = FALSE)
```
2. Bootstrap - Resampling
```{r}
set.seed(33834)
finalM2 <- train(classe ~ ., method = "rf", data = cleanTrain,prox = TRUE)
finalM2
confusionMatrix(finalM2)
#save(finalM1,file="finalM2.Rda")
varImp(finalM2, scale = FALSE)
```

Almost similar Variable Importance plots for both Bootstrapped & CV 

### Predicting against turning sample cleanTest using finalM1 & finalM2

```{r}
set.seed(12345)
predfinalM1 <- predict(finalM1,cleanTest,type="raw")
cleanTest$predRightM1 <- predfinalM1==cleanTest$classe
table(predfinalM1,cleanTest$classe)

predfinalM2 <- predict(finalM2,cleanTest,type="raw")
cleanTest$predRight <- predfinalM2==cleanTest$classe
table(predfinalM2,cleanTest$classe)
```
  
### Calculating Out Of Sample Estimation error

```{r}
missClass = function(values, prediction) {
    sum(prediction != values)/length(values)
}

# Error Rate with Model predfinalM1
errRate_M1 = missClass(cleanTest$classe, predfinalM1)
errRate_M1

# Error Rate with Model predfinalM2
errRate_M2 = missClass(cleanTest$classe, predfinalM2)
errRate_M2
```
  
### Predicting against holdout sample
```{r}
predfinalML_H <- predict(finalM1,sensorTest,type="raw")
```


